---
title: "Time Series Analysis: Pharmaceutical, cosmetic and toiletry goods retailing in Australia"
author: "Asghar Mustafa"
date: "31 May 2019"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This work aims to develop an optimal time series model for predicting future monthly turnover of retail of Pharmaceutical, cosmetic and toiletry goods in Australia. This would cover broadly 3 aspects of time series models i.e. Classical Methods, ETS Methods and ARIMA methods.  

## Data Reading

Setting up the libraries...

```{r libraries, echo=TRUE, message=FALSE}
library(fpp2)
library(readxl)
library(dplyr)
library(kableExtra)
library(knitr)

```

```{r echo=TRUE, message=FALSE}
x <- read_xlsx(here::here("/RetailDataIndividualFull.xlsx"), skip=3) 
y<- pull(x,"28905644")
```

Plotting the total Monthly Turnover (in millions AUD) from retail of Pharmaceutical, cosmetic and toiletry goods in Australia between April 1982 and December 2016.

```{r Data Plot}
myts<- ts(y,start = c(1982,4), frequency = 12)
train <- window(myts,end=time(myts)[length(myts)-24])
autoplot(myts,xlab="Time (month-Year)",ylab="Turnover (million AUD)",main="Monthly Retail Turnover")


```

The plot shows an upward trend from 1982 to 2016 with strong seasonal (shown by regular spikes) and cyclic component (shown by the repetitive shape/behaviour) in the data. 

## Classical Methods

Since it can be observed that the time series is heteroskedastic (no constant error variance), therefore it can be useful to transform the data. To do this box cox function in R was used with different lambda values including lambda=0 (log transformation), lambda=1/3 (cube root transformation), lambda=-1 (Inverse transformation) and lambda=0.10788 (box cox calculated transformation).
```{r Box Cox Transform}
lam=BoxCox.lambda(myts)
autoplot(BoxCox(myts,lambda=lam),xlab="Time (month-Year)",ylab="Turnover", main="BoxCox Transformation of monthly retail turnover")

```

As can be seen from the attached figure, the lambda=0.107 produces the most homoscedastic data and somewhat balances the seasonal and random component across the time series. Other transformations still show signs of hetroskedasticity across the time series.

```{r Train test split}
training<-window(myts, start=c(1982,4), end=c(2014,12))
test<-window(myts, start=c(2015,1))

autoplot(myts,xlab="Time (month-Year)",ylab="Turnover (million AUD)", main="Monthly retail turnover (Train & Test)") + 
  autolayer(training, series="Training") +
  autolayer(test, series="Test")

```

Moreover, the plot shows segregation of data in train and test.

```{r}
f1 <- snaive(training, h=length(test))

accuracy(f1,test)%>%kable()

```

```{r}
f2 <- rwf(training, drift=TRUE, h=length(test))
accuracy(f2,test)%>%kable()


```

```{r}
f3 <- naive(training, h=length(test))
accuracy(f3,test)%>%kable()

```

```{r Meanf}
f4 <- meanf(training, h=length(test))
accuracy(f4,test)%>%kable()
```

As can be seen from the above table, RMSE of seasonal naive is better than RMSE of naive. Moreover, snaive also has a better MAE, MAPE and MASE which suggests that Seasonal naive is a better forecasting method for future predictions of Monthly turnover for retail of Pharmaceutical, cosmetic and toiletry goods in Australia.

## ETS Models

As can be seen from the retail data has a linearly upward trending curve therefore this is considered an additive trend. The size of the variation (variance) in the turnover for the given time series also significantly increases with time which suggests the multiplicative seasonality of the time series. This multiplicative increase in seasonality along with additive increase in trend suggests that the model with a multiplicative error would give the best model. Therefore, the selected model is MAM i.e. multiplicative errors, Additive trend and Multiplicative Seasonality.  

```{r MAM}
fit_Assign3 <- ets(myts,model = 'MAM')
plot(fit_Assign3)
summary(fit_Assign3)

```

```{r MAM residuals}
checkresiduals(fit_Assign3)

```

As indicated by the ACF plot there are multiple lag plots which are significant. Moreover, the Ljung Box test also shows that the residuals are significant at any level of significance and thus rejects the NULL (p value approx 0).
As can be seen from residuals graph there is no major skewness in the residual distribution,although the auto correlation test is significant which means prediction intervals are not reliable, thus I will go with this model for point forecasts.

```{r MAM vs MAdM}
fit2_new<-ets(myts,model = 'MAM',damped=T)
summary(fit2_new)
checkresiduals(fit2_new)
f2_1<-forecast(fit2_new,h=24)
f2_11<-forecast(fit_Assign3,h=24)
autoplot(window(myts,start=c(2014,1)),xlab="Time (month-Year)",ylab="Turnover (million AUD)", main="Monthly retail turnover (Forecast)") + 
  autolayer(f2_1, series="MAdM")
autoplot(window(myts,start=c(2014,1)),xlab="Time (month-Year)",ylab="Turnover (million AUD)", main="Monthly retail turnover (Forecast)") + 
  autolayer(f2_11, series="MAM")


```

As shown by the graph the damped model has slightly higher prediction interval than the non damped model. On the other hand the forecasts from MAM are slightly higher compared to MAdM model because of the damping factor introduced by MAdM however the difference is very small due to the damping factor (phi) of 0.98 in the damped model thus both models look plausible.

## ARIMA Models

Now testing with ARIMA Models

```{r Box Cox Transformation}
lam=BoxCox.lambda(train)
autoplot(BoxCox(train,lambda=lam),xlab="Time (month-Year)",ylab="Turnover", main="BoxCox Transformation of monthly retail turnover")
```

Variance Stabilization: In the first step box cox transformation was chosen to remove heteroskedasticity from the data. This makes the turnover data homoscedastic as shown above.

```{r Seasonal difference}
myts_new<-BoxCox(train,lambda=lam)
myts_sta1<-diff(myts_new,lag=12)
autoplot(myts_sta1,xlab="Time (month-Year)",ylab=" Change in Turnover", main="Differencing 1 monthly retail turnover")

```

First order seasonal Difference: The attached figures shows first order difference of the transformed data at Lag 12. It shows slight evidence of the sinusoidal behaviour.


```{r Lag 1 difference}
myts_sta2<-diff(myts_sta1,lag =1)
autoplot(myts_sta2,xlab="Time (month-Year)",ylab=" Change in Turnover", main="Further Lag 1 monthly retail turnover")

```

Further lag 1 difference: A further lag 1 difference has been to remove the slight evidence of sinusoidal behaviour therefore the figure 3 now shows data to be highly stationary with no pattern of change.

```{r}
myts_sta2 %>% ggAcf()
myts_sta2 %>% pacf()
```

Since transformation involved 1 seasonal difference and 1 lagged difference therefor d=1 and ds=1. From the PACF it can be seen there are 2 significant successive lags and 1 significant seasonal lag so p=2 and ps=1. For ACF, there is 1 significant successive lag (second being just significant) and 2 consecutive seasonal lags therefore Q=1 and Qs=2. Thus suggested model is Arima (2,1,1)(1,1,2).

```{r ARIMA (2,1,1)(1,1,2)}
fit1<-Arima(train ,order=c(2,1,1),seasonal=c(1,1,2),lambda=lam)
fit1 %>% residuals() %>% ggtsdisplay(main="resids from Arima (2,1,1)(1,1,2)")
```

Trying out a few more ARIMA Models

```{r ARIMA (2,1,5)(2,1,2)}
Arima(train ,order=c(2,1,5),seasonal=c(2,1,2),lambda=lam)%>% residuals() %>% ggtsdisplay(main="Resids from Arima(2,1,5)(2,1,2)")

```

```{r ARIMA (3,0,1)(0,1,1)}
Arima(train ,order=c(3,0,1),seasonal=c(0,1,1),lambda=lam)%>% residuals() %>% ggtsdisplay(main="Resids from Arima(3,0,1)(0,1,1)")
```

Checking with Auto ARIMA
```{r Auto ARIMA}
auto.arima(train,lambda=lam)%>% residuals() %>% ggtsdisplay(main="Resids from Auto Arima")

```


```{r}
getrmse <- function(x,h,...)
{
  train.end <- time(x)[length(x)-h]
  test.start <- time(x)[length(x)-h+1]
  train <- window(x,end=train.end)
  test <- window(x,start=test.start)
  fit <- Arima(train,...)
  fc<-forecast(fit,h=h)
  return(round(accuracy(fc,test)[2,"RMSE"],4))
}
getrmse(train,h=24,order=c(2,1,5),seasonal=c(2,1,2),lambda=lam)
getrmse(train,h=24,order=c(3,0,1),seasonal=c(0,1,1),lambda=lam)
getrmse(train,h=24,order=c(2,1,1),seasonal=c(1,1,2),lambda=lam)
```

Using last 2 years data as the test set and the RMSE method. The RMSE is compared with few other arima methods and my chosen midel is ARIMA(2,1,5)(2,1,2). 

## Final Model

```{r ARIMA, ETS, Seasonal Naive compare}
fit_choose1<-Arima(train ,order=c(2,1,5),seasonal=c(2,1,2),lambda=lam)
fit_choose2<-ets(train,model = 'MAM')
fit_benchmark<- snaive(train, h=24,lambda=lam)

autoplot(window(myts,start=c(2015,1)),xlab="Time (month-Year)",ylab="Turnover(million$)", main="Forecasts 2017-18 with 3 ") +
  autolayer(forecast(fit_choose2,h=24),PI=FALSE,series="ETS MAM")+
  autolayer(fit_benchmark,PI=FALSE,series="S Naive")+
  autolayer(forecast(fit_choose1,h=24),PI=FALSE,series="ARIMA")
```

As shown the ARIMA Model most closely replicates the original time series data. Therefore my chosen model for retail data is the ARIMA model.

